\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-11-07 k-planar.tex
Minor touch-ups

  Consider the subgraph $H_v:=H[\{x\in V(T):\text{$v$ participates in $x$}\}]$.  Since $v\in Y_{x_0}$, $v$ participates in $x_0$, so $x_0\in V(H_v)$.  We claim that, for each $x\in V(H_v)\setminus\{x_0\}$, $H_v$ contains a path $y_0,\ldots,y_s$ of length $s\le k-1$ from $y_0:=x_0$ to $y_s:=x$.  Indeed, since $v$ participates in $x\neq x_0$, $\mathcal{S}$ contains a shortcut $P$ with $v\in V(P)$ and that crosses $x$.  Therefore $P$ contains a path $v_0,\ldots,v_s$ from $v_0:=v$ to some $v_s\in Y_x$.   Furthermore $v_s$ is an internal vertex of $P$ and $P$ has length at most $k$, so $s\le k-1$.  For each $i\in\{0,\ldots,s\}$, let $Y_{y_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.  For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or $y_{i-1}y_i\in E(H)$.
  Therefore $y_0,\ldots,y_s$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.} in $H_v$ of length at most $s\le k-1$.  This lazy walk contains a path from $y_0$ to $y_s$ of length at most $s\le k-1$, as promised.

@@

  Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.  Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$. Since $x_r\in S_{a}$, $a(v_r)=a$.  By \cref{short_path_in_h}, $H$ contains a path $y_0,\ldots,y_s$ from $y_0:=x_r$ to $y_s:=a$ of length at most $k-1$ such that $v_r$ participates in each of $y_0,\ldots,y_s$.  We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.   There are three cases to consider, depending on the location of $x_r$ relative to $x'$.

@@

    \item $x_r$ is a strict $T$-descendant of $x$.
      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$. Therefore $xy_i$ is an edge of $H^{+}$.  Therefore $x,y_i,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_i,\ldots,y_s$.

  \end{enumerate}
  In each case, we obtain a path $z_0,\ldots,z_{s'}$ in $H$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.  By definition, $a$ is a $T$-ancestor of $x$.  Since $v_r$ participates in each of $z_1,\ldots,z_{s'}$, $a=a(v_r)$ is also a $T$-ancestor of each of $z_1,\ldots,z_{s'}$.  Let $W$ be the sequence obtained from $z_0,\ldots,z_{s'}$ by replacing $z_i$ with the lowest common $T$-ancestor of $z_0,\ldots,z_i$, for each $i\in\{0,\ldots,s'\}$.  Then $W$ is a lazy walk in $H^+$ of length at most $k$ that begins at $z_0=x$ and ends at $z_{s'}=a$. By removing duplicates from $W$ we obtain a path in $H^+$ that is also a directed path in $\overrightarrow{H}^+$.  This completes the proof of \cref{general-bag-size}.
\end{proof}

>>>>

  Consider the subgraph $H_v:=H[\{x\in V(T):\text{$v$ participates in $x$}\}]$.  Since $v\in Y_{x_0}$, $v$ participates in $x_0$, so $x_0\in V(H_v)$.  We claim that, for each $x\in V(H_v)\setminus\{x_0\}$, $H_v$ contains a path $y_0,\ldots,y_s$ of length $s\le k-1$ from $y_0:=x_0$ to $y_s:=x$.  Indeed, since $v$ participates in $x\neq x_0$, $\mathcal{S}$ contains a shortcut $P$ with $v\in V(P)$ and that crosses $x$.  Therefore $P$ contains a path $v_0,\ldots,v_s$ from $v_0:=v$ to some $v_s\in Y_x$.   Furthermore $v_s$ is an internal vertex of $P$ and $P$ has length at most $k$, so $s\le k-1$.  For each $i\in\{0,\ldots,s\}$, let $Y_{y_i}$ be the part of $\mathcal{Y}$ that contains $v_i$.  For each $i\in\{1,\ldots,s\}$, the existence of $v_{i-1}v_{i}\in E(G)$ implies that $y_{i-1}=y_i$ or that $y_{i-1}y_i\in E(H)$.  Therefore $y_0,\ldots,y_s$ is a lazy walk\footnote{A \defin{lazy walk} in a graph $G$ is a walk in the pseudograph obtained by adding a self-loop at each vertex of $G$.} in $H_v$ of length at most $s\le k-1$.  This lazy walk contains a path from $y_0$ to $y_s$ of length at most $s\le k-1$, as promised.

@@

  Since $a\in C_x$, there exists a $T$-descendant $x'$ of $x$ such that $ax'\in E(J)$.  Therefore, there exists a shortcut $P:=v_0,\ldots,v_r$ in $\SS$ with $v_0\in S_{x'}$ and $v_r\in S_a$.  For each $i\in\{0,\ldots,r\}$, let $Y_{x_i}$ be the part of $\mathcal{Y}$ that contains $v_i$. Since $x_r\in S_{a}$, $a(v_r)=a$.  By \cref{short_path_in_h}, $H$ contains a path $y_0,\ldots,y_s$ from $y_0:=x_r$ to $y_s:=a$ of length at most $k-1$ such that $v_r$ participates in each of $y_0,\ldots,y_s$.  We will use this to show that $H$ contains a path $z_0,\ldots,z_{s'}$ of length $s'\le s$, from $z_0:=x$ to $z_{s'}:=a$, and such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.   There are three cases to consider, depending on the location of $x_r$ relative to $x'$.

  % \begin{figure}
  %   \begin{center}
  %     \begin{tabular}{ccc}
  %       % (1) & (2) & (3) \\
  %       \includegraphics[page=2]{figs/width} &
  %       \includegraphics[page=3]{figs/width} &
  %       \includegraphics[page=4]{figs/width}
  %     \end{tabular}
  %   \end{center}
  %   \caption{Three cases in the proof of \cref{general-bag-size}}
  %   \label{width_figure}
  % \end{figure}

@@

    \item $x_r$ is a strict $T$-descendant of $x$.
      Since  $y_0=x_r$ is a strict $T$-descendant of $x$ and $y_s$ is a $T$-ancestor of $x$, the path in $T$ from $y_s$ to $y_0$ includes $x$.  Since $H[\{y_0,\ldots,y_s\}]$ is connected, at least one of $y_0,\ldots,y_s$, say $y_i$, is contained in $B_x$. Therefore $xy_i$ is an edge of $H^{+}$.  Therefore $x,y_i,\ldots,y_s$ is a path in $H^+$ from $x$ to $a$ of length at most $s+1\le k$ and $v_r$ participates in each of $y_i,\ldots,y_s$.
  \end{enumerate}
  In each case, we obtain a path $z_0,\ldots,z_{s'}$ in $H$ of length at most $k$ from $z_0:=x$ to $z_{s'}:=a$ such that $v_r$ participates in each of $z_1,\ldots,z_{s'}$.  By definition, $a$ is a $T$-ancestor of $x$.  Since $v_r$ participates in each of $z_1,\ldots,z_{s'}$, $a=a(v_r)$ is also a $T$-ancestor of each of $z_1,\ldots,z_{s'}$.  Let $W$ be the sequence obtained from $z_0,\ldots,z_{s'}$ by replacing $z_i$ with the lowest common $T$-ancestor of $z_0,\ldots,z_i$, for each $i\in\{0,\ldots,s'\}$.  Then $W$ is a lazy walk in $H^+$ of length at most $k$ that begins at $z_0=x$ and ends at $z_{s'}=a$. By removing duplicates from $W$ we obtain a path in $H^+$ that is also a directed path in $\overrightarrow{H}^+$.  This completes the proof of \cref{general-bag-size}.
\end{proof}
